\vspace{-2cm} {\bf Online Appendix} \\ Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States

### 2019-05-24 paper/manuscript_paNote_OA.tex
splitting and revising OA

%============Article Title, Authors==================
\title{\vspace{-2cm} Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States}

@@

A local government's website is an important source of information about policies and procedures for residents, community stakeholders and scholars. Existing research in public administration, public policy, and political science has relied on manual methods of website content collection and processing, limiting the scale and scope of website content analysis. We develop a methodological pipeline that researchers can follow in order to gather, process, and analyze website content. Our approach, which represents a considerable improvement in scalability, involves downloading the entire contents of a website, extracting the text and discarding redundant information through a new method of boilerplate removal. We illustrate our methodological pipeline through the collection and analysis of a new and innovative dataset---the websites of over two hundred municipal governments in the United States. We build upon recent research that analyzes how variation in the partisan control of government relates to content made available on the government's website. Using a structural topic model, we find that cities with Democratic mayors provide more information on policy deliberation and crime control, whereas Republicans prioritize basic utilities and services such as water, electricity and fire safety.
%\noindent We explore the effect of transitions of power in municipal governments on the content of their websites. We hypothesize that when party control changes, city administrators modify the contents of their websites in order to fit the agenda of the new incumbent. To test this theory, we study cities in Indiana and Louisiana, two states in which all municipal elections are partisan and the parties of the candidates appear on the ballots. Snapshots of websites before and after transitions of power are acquired through the Wayback Machine. We apply statistical topic models based in latent dirichlet allocation, focusing on changes to the websites. We present results on both which topics see the greatest degree of change associated with transitions in city administrations, and how the topics modified differ with regard to political parties.

\end{abstract}
\thispagestyle{empty}
% \doublespacing
% Description of the possible challenges
\doublespacing
\section{Introduction}

Local governments convey voluminous information about all aspects of their policymaking, policy implementation, and public deliberation, via their official websites. The vital role of official websites in connecting the government and the governed has motivated a wave of research on the contents of government websites \citep[e.g.,][]{grimmelikhuijsen2010transparency,wang2005evaluating,osman2014cobra}. Despite the potential for automated scraping of website contents, the conventional approach to data collection in projects focused on government websites involves manual content extraction from each website in the dataset. Though highly accurate, the manual approach to data collection is costly, and cannot be scaled to capture even a fraction of the volume of content available on government websites. In this paper, we present a methodological pipeline that can be used to automatically scrape government websites in order to build datasets that can be used for text analysis. We provide an illustrative application in which we explore the ways in which the textual contents on city government websites in six American states (IN, LA, NY, WA, CA and TX) correlate with the partisanship of the city mayor.

Though there exists a variety of software tools that are designed to automatically scrape all of the files available at a website \citep{glez2013web}, raw website downloads have to be processed significantly before the files are adequately prepared for content (e.g., text, image) analysis. We describe and provide solutions to two central challenges in automatically gathering and analyzing website textual contents. First, plain text must be extracted from the files. This involves purging the files of syntax in HTML or other markup and programming languages and discarding any character encoding errors that result from reading in files with incorrectly specified character encodings. This challenge would arise in any context in which researchers sought to study the textual contents of websites, and is not unique to comparative analysis of government websites. The second challenge we address in our methodological pipeline is, however, specific to the research objective of comparing websites on the basis of a common lexicon. For any two governments, the textual signatures that most dramatically differentiate the textual contents of their websites consist of what we can call ``boilerplate'' text---header, footer, or other titling text that is designed to identify the website as being associated with a specific government entity (e.g., ``Welcome to the city of Santa Cruz'', ``The City of Los Angeles welcomes you''). This boilerplate text is replicated across many files that are associated with a government's website, but it provides little information regarding the form and/or function of the government. The second methodological innovation we offer in our pipeline is designed to minimize the impact of this boilerplate text on the comparative analysis of government website content.

Government websites provide information about how public policies shape the lives of local residents, and how local residents can engage with government to shape public policy. As such, government websites reflect both the results of and inputs to, the political leadership in the city. In our illustrative application, we explore the ways in which the contents of city government websites differ on the basis of the partisanship of the city's elected executive. A substantial body of research has found that the partisanship of the mayor affects city governance along multiple dimensions, including city budget priorities \citep{gerber2011mayors,de2016mayoral}, policies affecting inequality in cities \citep{einstein2016mayors}, and framing of criminal justice policy \citep{marion2013mayor}. Furthermore, recent media coverage of changes to government websites that follow transitions in party control suggest that changes in web content are salient government actions, as perceived by the general public \citep{sharfstein2017science,kirby2017trump,duarte2017deniable}. We study whether significant differences between city governments based on mayoral partisanship are reflected in the contents of city websites.

\section{The Significance of Government Website Content}

 According to \cite{Mayhew1974}, politicians engage in advertising, credit claiming and position taking in order to get re-elected. Official city websites allow mayors to perform all three of these functions. Their offices frequently take a prominent position on the front page, and many websites also feature a picture of the mayor. We present an example of this in Figure \ref{fig:eriemayor}. The Erie, Pennsylvania website homepage presents an image of the Democratic mayor, Joseph Schember, along with a list of laudable attributes of the city. In local politics, where campaign funds are low, this lends incumbents a crucial advantage in becoming more well-known among their constituencies \citep{stanyer2008elected}. Furthermore, municipal politics gives incumbents clear and tangible achievements they can point to, such as completed infrastructure projects, the acquisition of federal or state funding, or the hosting of city-wide events. City websites present an opportunity for local officials to brandish these accomplishments. Finally, they also give mayors a platform from which they can advertise their political beliefs. On municipal websites, this may not manifest in the form of brazen partisanship, but more subtle avenues are available. As noted by \cite{einstein2016mayors}, there are stark differences in the spending preferences of Democratic and Republican mayors. City websites can then be used to communicate the stance of a mayor on social or economic programs. Another advantage of websites with regard to communication is that unlike direct social interactions, officials have full control over them.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figures/eriemayor}
\caption{Screenshot from the homepage at \url{http://www.erie.pa.us/}, accessed on 06/14/2018. Image depicts  Democratic mayor of Erie, PA, Joseph Schember.}
\label{fig:eriemayor}
\end{figure}

Members of the public visit municipal government websites for a wide variety of purposes \citep{sandoval2012government}, and with significant regularity. In a survey conducted among a random sample of citizens in the state of Georgia in 2000---nearly two decades ago---found that 25\% of internet users reported visiting a local government website in the previous twelve months \citep{thomas2003new}. Furthermore, the use of a local government website is associated with an individual's perspective on government. \citet{tolbert2006effects} find that users of local government websites are more likely to trust local governments, and hold other positive attitudes related to local and federal governments. Lastly, in a study of residents of Kansas City, Missouri, \citet{ho2017government} find that participants' perceived quality of the city website is strongly associated with their perceptions of the overall effectiveness of the City's communication with the public.

The existing research that uses scraped websites provides an indication of the theoretical value of empirical analysis of web contents produced by governments, public officials, and candidates for office. The most pertinent literature to our research is the e-governance literature, which focuses on the online presence of governments from a usability and public service point of view. For the most part, research in this category develops a classification scheme to rate websites in terms of accessibility, ease-of-use, and function, and then hand-codes a set of websites according to these criteria \citep[e.g., ][]{Urban2002,Armstrong2011,Feeney2017}. As an example, \cite{grimmelikhuijsen2012developing} study local government websites with the goal of uncovering how they aid the goal of transparency. To this end, they analyze a set of Dutch municipalities in which air quality had deteriorated. The authors test whether local governments provide citizens with information about potential complications and solutions associated with this issue. Like most e-government studies, however, this publication does not make any use of automated text analysis. Branching out of the topic of e-government service provision, \citet{mcnutt2010virtual} uses Canadian government websites to study how well-connected government websites are to non-profits and intergovernmental organizations in different public policy domains.
%Websites have also played a major role in the field of media studies, as scholars have scraped and analyzed the online presence of newspapers, as well as the more diffuse world of online political blogs \citep{Adamic2005,Gentzkow2010}. \cite{Lin2011} provide a good example of a study which makes extensive use of automated content analysis - a necessity arising from its dataset of 66830 blog posts and 57221 online news articles. The authors estimate the political slant of these entities by counting the frequencies with which politicians of either side are mentioned and determine that blogs are generally more biased. Unfortunately for us, the authors don't go into the details of their text analysis, and offer no information on the acquisition and pre-processing of the data.
%Another well-known example fitting into this area of study is the set of studies conducted by King et al. \citep{KING2013,King2014,KING2017}, in which the authors study censorship by the country's government on its lively blogosphere. However, the authors also provide no information on how their data was collected ``our extensive engineering effort, which we do not detail here for obvious reasons [...]''.

The websites of politicians and their parties have also been the object of research. Researchers have found that in order to identify the constituencies, motives, and modes of communication among these actors, their websites can be very illuminating sources of information \citep{Druckman2009,Druckman2010,Cryer2017,Esterling2011,Esterling2011a,Norris2003,Therriault2010}. \cite{Druckman2009,Druckman2010} rely on the \textit{National Journal} to find the websites, then hand-coded them. \cite{Cryer2017} mentions that she relied on Archive-it, a web service of the Internet Archive. Though the Internet Archive may provide extensive coverage of high profile officials or national governments, we found that its coverage of municipal government websites was sparse and irregular.
\cite{Esterling2011,Esterling2011a} rely on hand-coded data by the Congressional Management Foundation, a nonprofit organization which aims to assist Congress. \cite{Therriault2010} use automated text analysis, and also has the most extensive overview of the associated methodology. However, the division of the website into sections (home page, topics, issues, details) is done by hand. The author acquired the websites from the Library of Congress (which only collected them from legislators who actually consented, and Therriault notes that this causes nonrandom missingness).

Important to our methodological objectives, research analyzing and improving the scraping, pre-processing and text analysis pipeline that is applicable to government websites is still in its infancy. \cite{Eschenfelder2002} provide something of an overview of how federal websites should be assessed from an e-governance point of view, but they largely focus on the substantive criteria that should be fulfilled, rather than the technical aspects of website acquisition and analysis. %In what follows we first define the target dataset---the textual contents of websites of United States municipalities, along with associated metadata on the municipalities and their governments. We then define a pipeline for data collection and analysis that includes methods to access government website URLs, scrape their raw contents from the World Wide Web, gather plain text from the website files, and identify boilerplate text within the plain text contents. Lastly, we illustrate the analysis of municipal government website text by exploring the relationship between the city mayors' party affiliations and the topical contents of the websites.

\section{Data: US Municipal Government Website Text}

For data availability reasons, we focus our analysis of municipal websites on six states---Indiana, Louisiana, New York, Washington, California, and Texas. The selection of states and cities is largely dictated by the presence of partisan mayors and availability of the relevant data. Municipal elections in Indiana and Louisiana are partisan across the board,\footnote{In Indiana, this includes only cities - incorporated municipalities with at least 2,000 inhabitants - as opposed to towns.} so our sample is primarily focused on these two states. Here, all cities with a website are included, resulting in a considerably larger sample than for the other four states. New York and Washington do not have nominally partisan elections, but for a subset of cities, partisanship can be determined through contribution data (see below for more detail on how we use this data). Finally, California and Texas contain a number of large cities, whose mayors are sufficiently well-known for their partisanship to be available. These six states also provide us with a sample that is well-balanced on a number of theoretically important indicators. One, each of the four geographic regions are represented with at least one state. Two, we have a fairly well-balanced sample with respect to the urban/rural cleavage, as both major cities and less densely populated areas are covered. Furthermore, the sample is politically balanced---we have three blue states (CA, WA, NY) and three red states (TX, IN, LA). Finally, our dataset contains some of the wealthiest states (NY, CA, WA and TX are \#2, \#8, \#9 and \#16 respectively, by GDP per capita \citep{BureauofEconomicAnalysis2017}), but also some of the poorer ones (IN and LA).

We acquired the website URLs from two sources: One, we scraped the URLs of city websites from their respective Wikipedia pages, which we found from lists of cities contained within each state. Two, the General Services Administration (GSA) maintains all `.gov' addresses, and provides a complete list of all such domains to the public.\footnote{The dataset is made available at \url{https://github.com/GSA/data/tree/gh-pages/dotgov-domains}. This list is updated once per month---we rely on the version released on January 16, 2017.} The data from the GSA contains the following variables: (1) domain name, specifically, the all-uppercase version of domain and top-level domain (for example, 'ABERDEENMD.GOV'); (2) the type of government entity to which the domain is registered, such as city, county, federal agency, etc; (3) for federal agencies, the name is specified; (4) the city in which the domain is registered. Naturally, the GSA's list does not contain cities which do not use a `.gov' website (or, in many cases, a city owns a registered `.gov' address, but uses a different one). Furthermore, some of the links are non-functional, and some of the county websites on the list are incorrectly marked as city websites (and vice versa). Since the GSA data is less complete and less reliable than the URLs found on Wikipedia, we mainly rely on the former and only supplement them with the GSA data if a specific city doesn't have a URL recorded on Wikipedia, or our tests (see below) find it to be non-functional.

@@

The partisanship of the mayor of each city is coded in different ways, depending on the state. For Indiana, where elections are nominally partisan, this information is accessible through the state government's website\footnote{\url{http://www.in.gov/apps/sos/election/general/general2015?page=office&countyID=1&officeID=32&districtID=-1&candidate=}}. For Louisiana, we received data on the outcomes of mayoral elections from the Local Elections in America Project (LEAP) \citep{marschall2013local}. For the other states, where mayoral elections are not nominally partisan (but the partisanship of the mayor is still well-known), we employed different means: For New York and Washington, we searched the state campaign finance websites, and coded the parties of the candidates based on the party committees from which they received donations. For California and Texas, where our data consists of highly populated cities, partisanship information was acquired from Ballotpedia\footnote{\url{https://ballotpedia.org/List_of_current_mayors_of_the_top_100_cities_in_the_United_States}}. Finally, we also scraped mayoral partisanship from the cities' Wikipedia pages. When compared to the other data sources above, (and manual searches in case of conflicts) Wikipedia proved to be very reliable and added additional cases to our dataset even for Indiana and Louisiana. Generally speaking, we found data scraped from Wikipedia, aided by manual corrections in case of missing or conflicting data, to be more reliable than data from governmental sources. The partisan breakdown of city websites is depicted in Table \ref{tab:cityparties}. The dataset has a relative party balance, with variation in each state.

@@

One of the more subtle aspects of local government is the presence of different types of government structures. Between council-manager governments and mayor-council governments \citep{morgan1992policy}---either in the weak or strong mayor variant \citep{desantis2002city}---there is a certain degree of variance in where a city's executive authority lies. Unfortunately, we do not have access to information about the type of governments across the breadth of our dataset and therefore cannot explore heterogeneity in the relationship between mayoral partisanship and municipal website contents based on the executive system variant. Given the prominent place that mayors tend to have on their cities' websites, we feel that any bias arising from this nuance should be minor. \cite{gerber2011mayors}, whose model is somewhat comparable to ours in the sense that they also test the effect of mayoral partisanship on city policy priorities (on a sample of large cities, for which this information is more easily available), find that the inclusion of this potential confounder does not affect the results. Information on other covariates (population and median household income - from the American Community Survey 5) was acquired through the API of the U.S. Census Bureau\footnote{\url{https://www.census.gov/data/developers/data-sets.html}}.

\section{The Web to Text Pipeline}

Once we have gathered the website files, we have the raw data necessary for text analysis, but it is not yet formatted effectively. As is common in text analysis applications, we need to pre-process our data \citep{denny2018text}. In this section, we describe our pre-processing pipeline, with which we take an archive of website files, and output a corpus of formatted plain text files that are suitable for comparative analysis with text as data methods. In this pipeline, we address two methodological challenges. First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step in the methodological pipeline is aimed simply at extracting clean plain text from this heterogeneous file base. The second step in our methodological pipeline is to process the text to remove boilerplate language---language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments. Some of the steps we take in this processing pipeline are universally applicable in the analysis of textual data, and some of them are most appropriate for the particular type of text analysis that we apply to this data---statistical topic modeling. We will clarify this distinction as we describe steps in our pipeline.

\subsection{Site to Text Conversion}
The format of a file has a major impact on whether and how textual data can be extracted from a document. For all text analysis projects, researchers need to consider file formats. For the most part, the file type of a document can be correctly determined through the filename ending---its extension. However, there are exceptions to this, which, if ignored, can lead to large amounts of improperly formatted text, arising from incorrectly converted documents, which leads to a general decrease in the amount of usable data. Two issues, in particular, need to be addressed: One, HTML files on city websites frequently do not have an ending but are still perfectly readable if correctly identified as such. Second, some documents contain the incorrect file ending. For example, we found thousands of documents that ended in .html, when they were actually PDFs. To accurately assess their type, we rely on the R package \texttt{wand} \citep{wand}, which is an R interface to the Unix library \texttt{libmagic} \citep{darwin2008libmagic}, which determines the type of a file on the basis of its file signature - or ``magic number''. This short sequence of bytes at the start (and sometimes end) of files is unique for each file type and therefore allows its correct identification through computer forensics tools such as \texttt{libmagic}.

Consequently, we rename all documents so that their file ending reflects their actual file type. This is strictly necessary because we rely on the \texttt{readtext} R package \citep{readtext}, which determines a document's type solely through its ending---to convert the files to plain text.\footnote{We have also experimented with several Unix-based alternatives, but found that they largely led to the same results as \texttt{readtext}.} The breakdown of the files by type is given in Table \ref{tab:filetypes}. The most frequent file types are HTML and PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

We then take several steps to pre-process the data as required for the subsequent analysis. Pre-processing choices should be contingent on the analysis being conducted with the text later and can have significant effects on the outcomes of an analysis \citep{denny2018text}. The type of text analysis we conduct---topic modeling---requires that the words in the document be meaningful and interpretable, and does not make use of the sequence of words within a document (i.e., is a bag-of-words method). %With the exception of file type conversion and boilerplate removal, the pre-processing steps taken here should therefore not be seen as universally applicable to all analyses of government website data.

The text documents are converted to UTF-8 character encoding and then stripped of dates, punctuation, numbers, and words connected by underscores. At this point, the documents of one city still closely resemble one another in the form of boilerplate content, be it website elements (i.e. "You are here", "Home", "Directory" etc.) in HTML documents, or commonly used forms or phrases in pdf, doc and docx files. This is an issue, because this boilerplate content causes the results of analyzing this data with text analysis methods to characterize documents primarily by the cities from which they originate (through their unique boilerplate structure, e.g. a menu with certain terms repeated on every site of the domain), and not the substantive features of their contents. Boilerplate removal is a useful step in many forms of text analysis, as the analysis is focused in on text that varies above and beyond a standard template for textual content. Our solution to this problem is described in more detail in Section \ref{boilerplate}.

The last round of preprocessing is intended to remove everything from the file that is not an English word. This step is tailored to our intention to use the text for topic modeling. The final preprocessing round includes setting every character to lowercase, as well as the removal of bullet points which frequently occur in HTML documents, extraneous whitespace, XML documents mislabeled as HTML files, and empty documents. Furthermore, some documents contain gibberish, often as a result of faulty or impartial optical character recognition applied to text that was produced through a non-machine-readable medium. To combat this problem, we employ two solutions. One, we use spellchecking, implemented through the \texttt{hunspell} R package \citep{hunspell}, to remove all non-English words.\footnote{Some of the cities, for example, Los Angeles, do contain a sizable proportion of Spanish content. The analysis of this content is beyond the scope of this paper but could be explored in future work, for example using methods of text processing that are applicable to multilingual corpora \citep{lucas2015computer}. } However, \texttt{hunspell} does not cover everything, either because some tokens are not actual words (for example artifacts from defective encoding), or because random sequences of characters just so happen to form words that exist in a dictionary (for example "eh" or "duh"). Since we rely on a bag-of-words model in which syntax does not matter, we can ameliorate these problems by removing all text except for whitespaces and the characters that appear in the English alphabet. Since a lot of the nonsensical text tends to be quite repetitive, we also delete all documents in which the proportion of unique to the total number of tokens is less than 0.15. Furthermore, \texttt{hunspell} does not spellcheck individual characters or two-character words, so we remove these token types entirely. Since these pre-processing steps reduce documents which are largely unsuitable to only a few tokens (i.e., word occurrences), we also remove all remaining documents containing less than 50 tokens. Finally, to remove words that are extremely rare (which also has the advantage of eliminating any remaining oddities) and thus add nothing substantive\footnote{Topic models essentially do not pick up on extremely rare words, so their inclusion is a waste of computational resources. Removing them in this manner is also the default preprocessing choice in the \texttt{stm} package.} to our models while increasing their computational cost, we also discard any token types that occur in only one document. We also conduct lemmatization to reduce words to their basic form. Lemmatization is similar to stemming but works in a somewhat more sophisticated manner by taking grammar and surrounding words into account to identify the dictionary form of a word. For example, the lemma of the word ``lemmatization'' would be ``lemmatize'', whereas most stemmers would simply chop off the ending, which would yield ``lemmatiz''. Thus, lemmatization makes the results more easily comprehensible. To this end, we rely on the R package \texttt{spacyr}, which provides an R implementation of the Python library \texttt{spaCy}.

\subsection{Boilerplate Removal}\label{boilerplate}
As noted above, city websites contain a large amount of text that is uninformative for its actual content, and therefore a hindrance to understanding through algorithmic text analysis. This is a common issue with textual data in which informative content is embedded in technically structured documents. See, e.g., \citet{burgess2016legislative,wilkerson2015tracing} and \citet{linder2018text} for examples of boilerplate removal in the analysis of legislative text. In the case of websites, lines in documents are generally quite informative, so all of our boilerplate removal efforts are done at the line level.

%this is outdated; not deleting it yet in case we need it for further reference
%Consequently we remove this content as follows: Each line of every document is compared to every line in every other document belonging to the same city. We count how many times each line is duplicated for that city. We remove any line occurring more than our chosen threshold of 10.\footnote{Empirically, lines tend to be duplicated either hundreds of times or only once or twice, if at all.} This means that each document only retains the information that is particular about it. We implement this algorithm through hash tables, which reduces the computational complexity from O($N^2$) to O($N$). Before this step is taken, we remove numbers and dates from the documents because they frequently make lines unique, despite the fact that they are virtually the same (for example different days on a city calendar).

\subsection*{Boilerplate Classification}
In order to determine whether a line should be discarded, we train a classifier on a human-coded sample. We sampled 500 lines from documents in each of the following five cities: Los Angeles, CA, Indianapolis, IN, New York, NY, Shreveport, LA, and Seattle, WA. To ensure that lines which occur more frequently in these cities (sometimes hundreds of thousands of times) had a higher probability of being scrutinized by the classifier, we use sampling weights equivalent to the proportion of total lines in a city's corpus made up by each specific line type. As an example, the most common line throughout all pages of the city of Seattle consists only of the word ``total'' and occurs 103,068 times. Similarly, the line ``page'' occurs 58,833 times. Even something completely nonsensical such as ``a a'' still appears on 376 occasions. To account for the higher likelihood of some lines being part of the training set, we use inverse probability weights in training the classifier---the weight of each line in the sample is 1/[number of occurrences in the corpus].\footnote{Note that the performance of the classifier is robust to the use of these weights and only changes by about one percentage point if they are not used.}

These 500 lines were then hand-coded as either substantively informative (210 lines) or not (290 lines). We then trained a number of different classifiers with this informativeness measure as the dependent variable. The independent variables we use are: (1) number of times the line was duplicated within the city, (2) the length of the line, in characters, (3) the number of tokens in the line, and (4) the median distance from the document midpoint to the position of the line itself. The purpose of these covariates is as follows:

\begin{itemize}
\item {\bf Line length:} The length of the line and the number of tokens are ways to find lines consisting of only a word or two. This is highly predictive of lines which are used as website headers and navigational elements, which are of zero substantive interest to us but are very effective at differentiating cities. These terms also happen to be fairly common, which causes them to be overweighted by the topic model.

\item {\bf Number of line duplications:} To directly address the latter problem, we include a measure of the number of times a line is duplicated within a city. Many lines occur hundreds or even thousands of times on a single website and therefore are terms that are highly predictive of the website, which causes the topic model to find topics that are highly predictive of cities, but not substantively informative.

\item {\bf Line position in the document} Since boilerplate terms such as navigational elements, headers, footers, and so on, should occur more frequently at the beginning and the end of websites, we attempt to identify such content as following: We measure the distance between the midpoint of a document and the position of a line, expressed as quantiles (to account for differing document lengths). Since lines can occur in multiple documents or multiple times in the same document, we take the median of these measures. Thus, for example, a line which often occurs at the beginning of documents might have a score of 0.45, whereas a line that tends to be found more in the center, and thus is indicative of more relevant content, might be scored with a 0.11 instead.
\end{itemize}

%Finally, the keyness measure: This indicator shows whether a term occurs disproportionally frequently in one document collection, compared to another. In our case, this would be one city's documents, compared to all others. Mathematically, this is a simple chi-square test, asserting whether the frequency of a word in a city is greater than its expected count, as determined by its count in all other documents, and the number of tokens in the city. The variable used in the model is the chi-square value of this test for each word (which varies and is therefore calculated individually for each city).\footnote{Or alternatively, the log of the absolute chi-square value, after which the previously negative values are made negative again.}

%A regression table for this model can be found in table 3. The associated accuracy in 5-fold cross-validation is 0.8325. The cross-validation accuracy of a simplified model with only the number of characters as a covariate is 0.825 (table 4). A model with the four covariates described above as well as interaction terms between each of them yields an accuracy of 0.885 (table 5).

%\input{tables/boilerplateClassifier1.tex}
\input{tables/boilerplateClassifierRFMetrics.tex}
%\input{tables/boilerplateClassifierRFImportance.tex}

We rely on a random forest as the final classifier, which offers slightly better performance than logistic regression.\footnote{We also tried SVM, boosted trees and AdaBoost, with similar results and chose the random forests because this method has a probabilistic basis and is more intuitive.} We assess the performance of this classifier through five-fold cross-validation. This means that the classifier is trained on 400 samples and then tested on the held-out set of 100, measuring metrics such as percent correctly predicted, precision, recall, and F1 score. This procedure is carried out five times so that each sample is part of the test set once. The aggregated (mean) results of this process can be found in Table \ref{randomForest}. For the implementation of this method, we rely on the R package \texttt{caret}, whose random forest classifier is based on the package \texttt{ranger}. We use this classifier to flag and remove all lines that are not classified (based on a threshold of $p = 0.5$) as substantively meaningful. The effect of this process on the corpus is illustrated with the corpus of Anchorage, AK (i.e. a city that isn't part of our sample used in the analysis) as an example in Figures \ref{boilerplate_before_after1} to \ref{boilerplate_before_after4}. Before the lines identified by the classifier as boilerplate are removed, lines with very few characters and words are the most common. After the removal, the distribution has changed---lines of medium length now occur more frequently than extremely short ones, which are unlikely to be substantively meaningful (see figures \ref{boilerplate_before_after1} and \ref{boilerplate_before_after2}). Furthermore, lines that are duplicated only a few times rather than dozens, hundreds or even thousands are now more common (see figure \ref{boilerplate_before_after3}). Finally, the position of the line in the documents is not as important to the random forest, and this also shows in the results. However, this feature still has a positive effect, as lines at either end of the document are a bit less common now (see figure \ref{boilerplate_before_after4}). Table \ref{tabBoilerplateIllustration} provides further illustration by listing the top 10 most likely boilerplate lines (in Anchorage, AK) -- all of which were flagged as such with a probability of 1. After all the preprocessing is set and done, our corpus consists of 259,099 documents.

>>>>

%============Article Title, Authors==================
\title{\vspace{-2cm} {\bf Online Appendix} \\ Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States}

@@

A local government's website is an important source of information about policies and procedures for residents, community stakeholders and scholars. Existing research in public administration, public policy, and political science has relied on manual methods of website content collection and processing, limiting the scale and scope of website content analysis. We develop a methodological pipeline that researchers can follow in order to gather, process, and analyze website content. Our approach, which represents a considerable improvement in scalability, involves downloading the entire contents of a website, extracting the text and discarding redundant information. We provide an R package that can be used to apply our proposed pipeline. We illustrate our methodological pipeline through the collection and analysis of a new and innovative dataset---the websites of over two hundred municipal governments in the United States. We build upon recent research that analyzes how variation in the partisan control of government relates to content made available on the government's website. Using a structural topic model, we find that cities with Democratic mayors provide more information on policy deliberation and crime control, whereas Republicans prioritize basic utilities and services such as water, electricity and fire safety.
%\noindent We explore the effect of transitions of power in municipal governments on the content of their websites. We hypothesize that when party control changes, city administrators modify the contents of their websites in order to fit the agenda of the new incumbent. To test this theory, we study cities in Indiana and Louisiana, two states in which all municipal elections are partisan and the parties of the candidates appear on the ballots. Snapshots of websites before and after transitions of power are acquired through the Wayback Machine. We apply statistical topic models based in latent dirichlet allocation, focusing on changes to the websites. We present results on both which topics see the greatest degree of change associated with transitions in city administrations, and how the topics modified differ with regard to political parties.

\end{abstract}
\thispagestyle{empty}
% \doublespacing
% Description of the possible challenges
\doublespacing

\section{Overview}

In this online appendix we include supporting information about our data, the data collection process, and some additional analyses. In the first section, we present additional details on data collection, along with some additional descriptive statistics. In the second section, we provide additional details on, and results from, our topic modeling analysis.

\section{Data collection methods and sources}

We acquired the municipal website URLs from two sources: One, we scraped the URLs of city websites from their respective Wikipedia pages, which we found from lists of cities contained within each state. Two, the General Services Administration (GSA) maintains all `.gov' addresses, and provides a complete list of all such domains to the public.\footnote{The dataset is made available at \url{https://github.com/GSA/data/tree/gh-pages/dotgov-domains}. This list is updated once per month---we rely on the version released on January 16, 2017.} The data from the GSA contains the following variables: (1) domain name, specifically, the all-uppercase version of domain and top-level domain (for example, 'ABERDEENMD.GOV'); (2) the type of government entity to which the domain is registered, such as city, county, federal agency, etc; (3) for federal agencies, the name is specified; (4) the city in which the domain is registered. Naturally, the GSA's list does not contain cities which do not use a `.gov' website (or, in many cases, a city owns a registered `.gov' address, but uses a different one). Furthermore, some of the links are non-functional, and some of the county websites on the list are incorrectly marked as city websites (and vice versa). Since the GSA data is less complete and less reliable than the URLs found on Wikipedia, we mainly rely on the latter and only supplement them with the GSA data if a specific city doesn't have a URL recorded on Wikipedia, or our tests (see below) find it to be non-functional.

@@

The partisanship of the mayor of each city is coded in different ways, depending on the state. For Indiana, where elections are nominally partisan, this information is accessible through the state government's website\footnote{\url{http://www.in.gov/apps/sos/election/general/general2015?page=office&countyID=1&officeID=32&districtID=-1&candidate=}}. For Louisiana, we received data on the outcomes of mayoral elections from the Local Elections in America Project (LEAP) \citep{marschall2013local}. For the other states, where mayoral elections are not nominally partisan (but the partisanship of the mayor is still well-known), we employed different means: For New York and Washington, we searched the state campaign finance websites, and coded the parties of the candidates based on the party committees from which they received donations. For California and Texas, where our data consists of highly populated cities, partisanship information was acquired from Ballotpedia\footnote{\url{https://ballotpedia.org/List_of_current_mayors_of_the_top_100_cities_in_the_United_States}}. Finally, we also scraped mayoral partisanship from the cities' Wikipedia pages. When compared to the other data sources above, (and manual searches in case of conflicts) Wikipedia proved to be very reliable and added additional cases to our dataset even for Indiana and Louisiana. Generally speaking, we found data scraped from Wikipedia, aided by manual corrections in case of missing or conflicting data, to be more reliable than data from governmental sources.\footnote{In Indiana, the data includes only cities - incorporated municipalities with at least 2,000 inhabitants - as opposed to towns.}

Information on other covariates (population and median household income - from the American Community Survey 5-Year Data (2015)) was acquired through the API of the U.S. Census Bureau\footnote{\url{https://www.census.gov/data/developers/data-sets.html}}.

%One of the more subtle aspects of local government is the presence of different types of government structures. Between council-manager governments and mayor-council governments \citep{morgan1992policy}---either in the weak or strong mayor variant \citep{desantis2002city}---there is variance in where a city's executive authority lies. We do not have access to information about the type of governments across the breadth of our dataset. Given the prominent place that mayors tend to have on their cities' websites, we feel that any bias arising from this nuance should be minor. \cite{gerber2011mayors}, whose theory is somewhat comparable to ours, find that the inclusion of this potential confounder does not affect the results.

Tables \ref{tab:cityparties} and \ref{tab:filetypes} provide additional information about the data collected for this project. In Table \ref{tab:cityparties}, we present the state-by-state breakdown of the mayoral partisanship of the cities collected in the respective state. In Table \ref{tab:filetypes}, we present the distribution of file extensions before and after processing.

@@

@@

\section{Supplemental Informational on Topic Modeling Application}

The structural topic model is implemented in the R package \texttt{STM} \citep{stm}. We use 60 topics---the number recommended by the authors\footnote{For this recommendation, see the documentation for the function \texttt{stm()} in version 1.3.0 of the R package \texttt{stm} \citep{stm}.} for medium- to large-sized corpora.\footnote{Since our corpus is at the larger end of that spectrum, we also estimated a model with 120 topics, but found no notable differences.} We use four covariates: First, \textit{party}, to estimate the difference in topic prevalence based on whether mayors are Republican or Democratic. Second, \textit{city population}, which the literature frequently emphasizes as a determinant of the issues a city faces (see, for example, \cite{Guillamon2013}). Third, we control for wealth by relying on \textit{median income} as a covariate, which we use as a proxy for the tax base in a city. Fourth and finally, we include state dummy variables, which should account for language that is associated with state-specific issues, and general background variables that vary across states.\footnote{The ``Fightin' Words'' methodology developed by \citet{Monroe2008} could also be used to analyze word-frequency differences between cities based on mayors' partisanship, but we elected to use the structural topic model since, unlike ``Fightin' Words'' , the structural topic model enables us to adjust for several other features through multiple regression.}

\subsection{Supplemental results}

In Tables \ref{tabSTMtopwords120_1} and \ref{tabSTMtopwords120_2} we present results of the STM with 120 topics organized according to the effect of mayoral partisanship on topic prevalence. The partisan themes in the 120-topic STM mirror those in the 60-topic model, with cities led by Democratic mayors focused disproportionately on finances (e.g., Topics 75, 71, 61) and social problems (e.g., Topics 93, 101, 39, 52), and cities led by Republican mayors focused disproportionately on basic services and utilities (e.g., Topics 114, 11,  86, 116).

In Tables \ref{tabSTMtopwords60_income} and \ref{tabSTMtopwords60_population} we present the topics ordered according to the effects of median income and population, respectively. Considering income-based variation in topics, there are several topics prevalent in more wealthy cities that focus on initiatives that go well-beyond standard city services---downtown and building revitalization (Topics 39 and 12), renewable energy (Topic 20), bike/pedestrian-oriented development (Topic 60), wildlife conservation (Topic 3). Such topics cannot be found among those that are more prevalent in less wealthy cities. When it comes to population, more populous cities deal disproportionately with issues related to public health (Topics  2 and 11), crime (Topic 59), homelessness (Topic 52), and diversity (Topic 6).

\input{tables/stmTopWords60_income.tex}
\input{tables/stmTopWords60_population.tex}
